---
title: "Correlated multivariate OU inference with kfl1ou"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Correlated multivariate OU inference with kfl1ou}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

This workflow separates three questions that should not be conflated:

1. where the adaptive optimum may have shifted;
2. whether evolutionary innovations are correlated among traits; and
3. how much uncertainty is introduced by model and shift selection.

Start by validating and, if desired, normalizing the data. The normalization
factor is retained so rate parameters can be converted back to the original
tree-time unit.

```{r, purl=FALSE}
library(kfl1ou)
data(lizard.tree, lizard.traits)

keep <- lizard.tree$tip.label[seq_len(30)]
tree <- ape::drop.tip(lizard.tree, setdiff(lizard.tree$tip.label, keep))
dat <- adjust_data(tree, lizard.traits[keep, 1:3])
```

Fit a full evolutionary innovation covariance only when there are enough
species and residual contrasts for the number of traits. A distinct alpha for
each trait is more flexible but also less identifiable.

```{r, purl=FALSE}
fit <- estimate_shift_configuration(
  dat$tree,
  dat$Y,
  max.nShifts = 2,
  criterion = "BIC",
  trait.covariance = "full",
  alpha.structure = "diagonal",
  likelihood.engine = "auto"
)

evolutionary_vcov(fit)
fit$tip.trait.correlation
original_time_parameters(fit)
```

Use parametric calibration for covariance comparison and repeat the complete
shift search when uncertainty from shift selection is scientifically relevant.

```{r, purl=FALSE}
comparison <- compare_trait_covariance(
  dat$tree, dat$Y, nboot = 200, seed = 1, nCores = 2
)

support <- l1ou_bootstrap_support(
  fit, nItrs = 200, type = "parametric", seed = 2,
  multicore = TRUE, nCores = 2
)
summarize_shift_uncertainty(support, dat$tree)

confint(
  fit, method = "parametric", selection = "full",
  interval = "basic", nsim = 200, seed = 3
)
```

Regularized covariance estimates can stabilize difficult fits, but their BIC
values are sensitivity scores evaluated at penalized estimates. Report both
regularized and unregularized results, optimizer boundaries, bootstrap failure
rates, tree/root-model sensitivity, and the candidate configurations actually
evaluated. Exact historical shift timing is not identifiable from extant-tip
data alone.
